\documentclass[12pt,a4paper]{article}

%\setcounter{secnumdepth}{1}

%opening
\title{Thesis Plans}
\author{Tobin South}

\begin{document}

\maketitle

\section{Title Ideas}

\section{Sections}
\subsection{Introduction}
Introduce the problem we are trying to solve. \\
What actually is the goal\\
\subsection{Background}
\subsubsection{Entropy Basics}
Basics and History, essentially a literature review on entropy as a topic.
\subsubsection{A introduction to Natural Language Processing}
Tokenizing, Def: BOW, Token Sequence, Time of Speaking
A small literature review on the work of tokenizing and different techniques.
\subsubsection{Process Entropy and Natural Language}
An introduction to the idea of language as a non-markovian process. A description of the core maths that underlies process entropy in stationary ergodic processes. Def: Stationary, Ergodic, Markvoian.
\subsection{Information Measures and Distances}
Predictability and Self-information\\
KL Divergence \\
JS Distance \\
Mutual information?
\subsection{Chapter 1}
\subsubsection{Data}
Twitter?\\
Describe Text and Provide Analysis\\
Allsides explained and evaluated with reliability. \\
\subsubsection{Basic Works}
Initial clustering of data and TFIDF \\
LDA? \\
Exploration of why this doesn't work.
\subsubsection{Entropy on this text}
Full entropy (not time based) \\
Plots of distributions, groupings by bias\\
signal to noise ratio
\subsection{Chapter 2}
\subsubsection{A background on networks}
 A background on networks \\
 directed graphs\\
 undirected graphs \\
 networks vs graphs \\
 dense graphs
 centrality measures
\subsubsection{Directed Networks}
Create a directed network, look for clustering and report metrics\\
more to come




\end{document}
