%!TEX root = ../thesis.tex
\chapter{News Media Twitter Accounts\label{sec:app_accounts}}

\section{Included News-Media Organisations}

\begin{center}
\begin{longtable}{p{45mm}llrr}
	\caption{All included news-media organisations in clean dataset with full names, Twitter handles, the number of Twitter account followers, the numbers of tweets for 2019, and the political bias assigned to the organisation by AllSides.} \label{tab:app_included_accounts} \\

	Name &  
	Twitter Account & 
	\begin{tabular}{@{}l@{}}Assigned  \\ Bias\end{tabular} 
	 &  
	\begin{tabular}{@{}l@{}}Number of \\ Tweets\end{tabular}
	&  Followers \\ \hline
	\endfirsthead
	\caption[]{All included news-media organisations in clean dataset with full names, Twitter handles, the number of Twitter account followers, the numbers of tweets for 2019, and the political bias assigned to the organisation by AllSides.}\\

	Name &  
	Twitter Account & 
	\begin{tabular}{@{}l@{}}Assigned  \\ Bias\end{tabular} 
	 &  
	\begin{tabular}{@{}l@{}}Number of \\ Tweets\end{tabular}
	&  Followers \\ \hline
	\endhead

	\hline \multicolumn{5}{r}{{Continued on next page}} \\ 
	\endfoot

	\hline 
	\endlastfoot

	\input{appendix1/figs/appendix_accounts}

\end{longtable}
\end{center}

\newpage

\section{Excluded News-Media Organisations}

\begin{center}
\begin{longtable}{p{42mm}llp{35mm}}
	\caption{A list of news-media organisations that are listed by AllSides but are not included in the cleaned data. The reason for removal from the dataset in listed.} \label{tab:app_removed_accounts} \\

	Name &  Twitter Account & \begin{tabular}{@{}l@{}}Assigned  \\ Bias\end{tabular}  &  \begin{tabular}{@{}l@{}}Reason for  \\ Removal\end{tabular}  \\ \hline
	\endfirsthead
	\caption[]{A list of news-media organisations that are listed by AllSides but are not included in the cleaned data. The reason for removal from the dataset in listed.} \\

	Name &  Twitter Account & \begin{tabular}{@{}l@{}}Assigned  \\ Bias\end{tabular}  &  \begin{tabular}{@{}l@{}}Reason for  \\ Removal\end{tabular}  \\ \hline
	\endhead

	\hline \multicolumn{4}{r}{{Continued on next page}} \\ 
	\endfoot

	\hline 
	\endlastfoot

	\input{appendix1/figs/removed_accounts}

\end{longtable}
\end{center}

\newpage

\section{News-Media Organisation Locations}

\begin{center}
\begin{longtable}{lrp{37mm}}
	\caption{The declared location and number of followers for each news-media organisation in the cleaned dataset. The declared location is the location as listed on the organisations Twitter account.} \label{tab:app_locations} \\

	Organisation Name &  Number of Followers &     Declared Location \\ \hline 
	\endfirsthead
	\caption[]{The declared location and number of followers for each news-media organisation in the cleaned dataset. The declared location is the location as listed on the organisations Twitter account.}  \\

	Organisation Name &  Number of Followers &     Declared Location \\ \hline 
	\endhead

	\hline \multicolumn{3}{r}{{Continued on next page}} \\ \hline
	\endfoot

	\hline 
	\endlastfoot

	\input{appendix1/figs/locations}

\end{longtable}
\end{center}

